\normalfont\spacedallcaps{Hazardous asteroids forecast via Markov random fields}

### 2021-11-24 Manuscript/manuscript.tex
PUB

Lets now move on the mixed interaction model: first we considered its implementation in the mgm package \cite{mgm,haslbeck2015mgm}. We choose as $k$ parameter 2 and a cross validation (CV) with ten folds. The result is reported in Fig. \ref{mgm}. From this figure we see that the features that are conditionally dependant to the Hazardous features are: the absolute magnitude, the min orbit intersection, and the eccentricity.  All these connections are supported by the theory: the min orbit intersection is the main parameter for the hazardous value, the eccentricity on the other side can be tough as a parameter that describes how near the celestial body moves at the perihelion (indeed this parameter is correctly linked with the perihelion distance) and the Absolute Magnitude is also a meaningful parameter for the hazard evaluation since, if the asteroid is too small it will be destroyed by the earth atmosphere. It is worth mentioning that this parameter is correctly not connected with all quantities related to the orbital parameter except for the mean anomaly. Furthermore the parameters related to the diameters are also not connected with the orbital parameters. Furthermore we see also that the mean velocity is not directly related to the mean motion but there are the orbital parameters within them as expected. Thus it can be said that the model produced is almost consistent with the astronomic laws. On this basis, we can evaluate the performances of the model in term of confusion matrix, ROC curve and $\phi$ parameters (we used for this purpose the Caret R package \cite{kuhn2008building,caret}). The confusion matrix obtained is reported in Fig. \ref{mgm_confusion}: the ROC curve is given in Fig. \ref{ROC_mgm} and it corresponds to a $\phi=0.6$. We will comment these performances in a few when we will present the performances of other algorithms for the same dataset. Before this we would show the graphical models obtained with different implementation of mixed interaction model: we considered the function \textit{mmod()} of gRim package \cite{hojsgaard2012graphical} and the \textit{minforest()} of gRapHD (which uses the minForest method) \cite{de2009high}. Both were obtained with a stepwise algorithm.  Their result are reported in Fig. \ref{mmod} and \ref{minforest}. We see that in these models the minimum diameter is linked with the hazordous feature, and this is correct, but what is without a physical meaning is that this quantity is also linked to the features connected with the orbital parameters. Indeed, while for the magnitude (as for mgm model) a weak relationship can be admitted since this depends also on the distance between the asteroids and the observer and from the orbit, a volume and thus a mass dependence is not acceptable. In principle one can set these link as forbidden in the algorithm (blacklist), but in the author view it is preferable a model were the result is obtained without boundaries, instead of a one obtained with a large number of boundaries. Thus since a correct model was obtained without the imposition a blacklist, the mgm, we simply reject the mmod and minforest ones in favour of the first one. As consequence this is the reason why we report the performances of mgm model only.

@@

From which we get that the velocity of the asteroid is maximum at the perihelion, and minimum at the aphelion. Their values that are equal respectively to \cite{murray1999solar}:

@@

The solar system is composed not only by planets and the Sun, but also, as shown in Fig. \ref{Orbit_asteorids}, by a plethora of small bodies\footnote{Small with respect to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}. In particular if the orbit of these items is near less then 1.3 AU such items are called near-Earth object (NEO). Such items are of three type: comets, meteoroids and asteroids. The first one are icy bodies that moving near the sun melts and release gases giving coloured tails. Meteroids are small \footnote{With respect to asteroids} rocky/metallic items with a diameter less than 1 meter. Asteroids, on the other side, are items with a diameter larger than 1 m \cite{burbine2016asteroids}. The first two items are excluded from the present analysis.  
The near asteroids, which are mostly located on the sites reported in Fig. \ref{heic1715c}, are classified, as shown in Fig. \ref{neo_orbit_types}, according to their semi-major axis (a), perihelion distance (q), and aphelion distance (Q). Here we report the CNEOS definition \cite{nasa_classification}:

@@

On the basis of this definition, the asteroids contained in the dataset here analysed are classified as hazardous or not.

>>>>

Lets now move on the mixed interaction model: first we considered its implementation in the mgm package \cite{mgm,haslbeck2015mgm}. We choose as $k$ parameter 2 and a cross validation (CV) with ten folds. The result is reported in Fig. \ref{mgm}. From this figure we see that the features that are conditionally dependant to the Hazardous features are: the absolute magnitude, the min orbit intersection, and the eccentricity.  All these connections are supported by the theory: the min orbit intersection is the main parameter for the hazardous value, the eccentricity on the other side can be tough as a parameter that describes how near the celestial body moves at the perihelion (indeed this parameter is correctly linked with the perihelion distance) and the Absolute Magnitude is also a meaningful parameter for the hazard evaluation since, if the asteroid is too small it will be destroyed by the earth atmosphere. It is worth mentioning that this parameter is correctly not connected with all quantities related to the orbital parameter except for the mean anomaly. Furthermore the parameters related to the diameters are also not connected with the orbital parameters. In addition we see also that the mean velocity is not directly related to the mean motion but there are the orbital parameters within them as expected. Thus it can be said that the model produced is almost consistent with the astronomic laws. On this basis, we can evaluate the performances of the model in term of confusion matrix, ROC curve and $\phi$ parameters (we used for this purpose the Caret R package \cite{kuhn2008building,caret}). The confusion matrix obtained is reported in Fig. \ref{mgm_confusion}: the ROC curve is given in Fig. \ref{ROC_mgm} and it corresponds to a $\phi=0.6$. We will comment these performances in a few when we will present the performances of other algorithms for the same dataset. Before this we would show the graphical models obtained with different implementation of mixed interaction model: we considered the function \textit{mmod()} of gRim package \cite{hojsgaard2012graphical} and the \textit{minforest()} of gRapHD (which uses the minForest method) \cite{de2009high}. Both were obtained with a stepwise algorithm.  Their result are reported in Fig. \ref{mmod} and \ref{minforest}. We see that in these models the minimum diameter is linked with the hazordous feature, and this is correct, but what is without a physical meaning is that this quantity is also linked to the features connected with the orbital parameters. Indeed, while for the magnitude (as for mgm model) a weak relationship can be admitted since this depends also on the distance between the asteroids and the observer and from the orbit, a volume and thus a mass dependence is not acceptable. In principle one can set these link as forbidden in the algorithm (blacklist), but in the author view it is preferable a model were the result is obtained without boundaries, instead of a one obtained with a large number of boundaries. Thus since a correct model was obtained without the imposition a blacklist, the mgm, we simply reject the mmod and minforest ones in favour of the first one. As consequence this is the reason why we report the performances of mgm model only.

@@

From which we get that the \textcolor{red}{velocity} \footnote{In the dataset this quanitity is called relative velocity because the it the reference system is Earth} of the asteroid is maximum at the perihelion, and minimum at the aphelion. Their values that are equal respectively to \cite{murray1999solar}:

@@

The solar system is composed not only by planets and the Sun, but also, as shown in Fig. \ref{Orbit_asteorids}, by a plethora of small bodies\footnote{Small with respect to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}. In particular, if the orbit of these items is near less then 1.3 AU such items are called near-Earth object (NEO). Such items are of three type: comets, meteoroids and asteroids. The first one are icy bodies that moving near the sun melts and release gases giving coloured tails. Meteroids are small \footnote{With respect to asteroids} rocky/metallic items with a diameter less than 1 meter. Asteroids, on the other side, are items with a diameter larger than 1 m \cite{burbine2016asteroids}. The first two bodies are excluded from the present analysis. The near asteroids, which are mostly located on the sites reported in Fig. \ref{heic1715c}, are classified, as shown in Fig. \ref{neo_orbit_types}, according to their semi-major axis (a), perihelion distance (q), and aphelion distance (Q). Here we report the CNEOS definition \cite{nasa_classification}:

@@

This is the formal definition by which the asteroids contained in the dataset here analysed are classified as hazardous or not.
